---
title: "Assigment_02"
author: "Hasan Palaz"
date: "October 21, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

######Q1
Create two matrices from the given set of numbers 
X1 = {2,3,7,1,6,2,3,5,1} and x2 = {3,2,9,0,7,8,5,8,2}

######A1
```{r}
# Define X1 and X2 datasets
X1 <- c(2,3,7,1,6,2,3,5,1)
X2 <- c(3,2,9,0,7,8,5,8,2)

# Create 3x3 matrix_01 and matrix_02 from X1 and X2 vectors
matrix_01 <- matrix(X1, ncol=3)
matrix_02 <- matrix(X2, ncol=3)

# Lets's look our new matrices
matrix_01
matrix_02

```

######Q2
Find the matrix product.

######A2
```{r}
# multiplication of matrix_01 and matrix_02
matrix_03 <- matrix_01 %*% matrix_02
matrix_03
```
######Q3
What does correlation coefficient represent? 
How do you find the correlation between 
(i) two categorical variables and 
(ii) categorical variable and continuous variable.

######A3
What does correlation coefficient represent?

A) In statistics, the correlation coefficient r measures the strength and direction of a linear relationship between two variables on a scatterplot. The value of r is always between +1 and –1.

B) The further an r value is from zero, the stronger the relationship between the two variables.

C) The sign of r indicates the nature of relationship. A positive r indicates a positive relationship, and a negative r indicates a negative relationship.

How do you find the correlation between two categorical variables?

A) Contingency Tables, two way table and stacked column plot are used to find the correlation between two categorical variables. 

How do you find the correlation between categorical variable and continuous variable?

A) Z-test/t-test and ANOVA are used to find the correlation between categorical variable and continuous variable. 

######Q4
Load the Animals data set from the “MASS” package

######A4
```{r}
# If the MASS package is not installed, firstly we do that:
#install.packages('MASS')
library(MASS)
data(Animals)
head(Animals)
```

######Q5
Find the correlation coefficient of brain and body in this data set and comment on the relation between them

######A5
```{r}
# Correlation coefficient of brain and body in Animals dataset
cor(Animals$body, Animals$brain)
```

Comment on the relation betweenbrain and body:

A) There is no relation (correlation) between brain and body in Animals dataset.
Because, the correlation of brain and body in Animals dataset is -0.00534, around ZERO.


######Q6
Load USArrests data set. Comment on the distribution of the variables. 
Is there any relation between UrbanPop and three crimes?

######A6
Load USArrests data set.
```{r}
# Load USArrest data set and check details
data(USArrests)
USArrests
names(USArrests)
head(USArrests)
dim(USArrests)
```
Comment on the distribution of the variables.
```{r}
boxplot(USArrests$Murder, horizontal=TRUE, col= c("red"),main="Box plot of Murder Rate")
boxplot(USArrests$Assault, horizontal=TRUE, col= c("yellow"),main="Box plot of Assault Rate")
boxplot(USArrests$Rape, horizontal=TRUE, col= c("blue"),main="Box plot of Rape Rate")
boxplot(USArrests$UrbanPop, horizontal=TRUE, col= c("brown"),main="Box plot of UrbanPop")
hist(USArrests$Murder, col="purple",main="Histogram of Murder Rate",xlab="Murder Rate")
hist(USArrests$Assault, col="green",main="Histogram of Assault Rate",xlab="Assault Rate")
hist(USArrests$Rape, col="pink",main="Histogram of Rape Rate",xlab="Rape Rate")
hist(USArrests$UrbanPop, col="gray",main="Histogram of UrbanPop",xlab="UrbanPop")
summary(USArrests)

```

Comment on the distribution of the variables.

1) Murder rate has a distribution with a very large positive skew. This distribution is not symmetric

2) Assault rate has the two distinct peaks: one at 100-150 and the other at 250-300. Its distribution is not symmetric.
This distribution with two peaks is called a bimodal distribution.

3) Rape rate has a positive skew. Its distribution is not symmetric. It is also described as "skewed to the right."

4) UrbanPop has an almost symmetric (normal) distribution.


Is there any relation between UrbanPop and three crimes?
Firstly, correlation coefficients of the USArrest are calculated
```{r}
# Correlation coefficient of the USArrests dataset
cor(USArrests)
```
Secondly, Scatter Plots of the USArrest should be reviewed.
```{r}
plot(USArrests$UrbanPop, USArrests$MUrder)
plot(USArrests$UrbanPop, USArrests$Assault)
plot(USArrests$UrbanPop, USArrests$Rape)
```

According to correlations and cattering plots, it can be said that;

1) There is no relation (r=0.06957262, because r<0.1) between UrbanPop and Murder.

2) There is a weak relation (r=0.25887170, because r<0.3) between UrbanPop and Assault. More observations are needed to find out relation between UrbanPop and Assault.

3) There is a moderate relation (r=0.41134124, because 0.3 < r < 0.5) between UrbanPop and Rape. According to scattering plot Rape against UrbanPop, there are some outliers, but there does seem to be a relationship between Rape and UrbanPop. There is an indication of a correlation and more analysis could be performed.


######Q7
Which states has most and least assault, murder, and rape arrests?

######A7

```{r}
# Most Assult State
Assult_max <- which(USArrests$Assault == max (USArrests$Assault))
print('Most Assult State:')
rownames(USArrests)[Assult_max]
# Least Assult State
Assult_min <- which(USArrests$Assault == min (USArrests$Assault))
print('Least Assult State:')
rownames(USArrests)[Assult_min]

# Most Murder State
Murder_max <- which(USArrests$Murder == max (USArrests$Murder))
print('Most Murder State:')
rownames(USArrests)[Murder_max]
# Least Murder State
Murder_min <- which(USArrests$Murder == min (USArrests$Murder))
print('Least Murder State:')
rownames(USArrests)[Murder_min]

# Most Rape State
Rape_max <- which(USArrests$Rape == max (USArrests$Rape))
print('Most Rape State:')
rownames(USArrests)[Rape_max]
# Least Rape State
Rape_min <- which(USArrests$Rape == min (USArrests$Rape))
print('Least Rape State:')
rownames(USArrests)[Rape_min]
```

######Q8
List the states which have assault arrests more than median of the country.

######A8

```{r}
# Assult arrest of median of the cuntry
assault_median = median(USArrests$Assault)
assault_median
# List the states which have assault arrests more than median of the country
subset(USArrests, Assault > assault_median, select= c(UrbanPop, Assault))
```


######Q9
Which states are in the bottom 25% of murder?

######A9

```{r}
# Bottom 25% of murder, lowQuartile
lower_quartille_murder_rate <- quantile(USArrests$Murder)[2]
lower_quartille_murder_rate 
# States are in the bottom 25% of murder
subset(USArrests, Murder < lower_quartille_murder_rate, select= c(UrbanPop, Murder))
```
######Q10
Which states are in the top 25% of the murder.

######A10

```{r}
# Top 25% of the murder
upper_quartile_murder_rate <- quantile(USArrests$Murder)[4]
upper_quartile_murder_rate
# States are in the top 25% of the murder
subset(USArrests, Murder > upper_quartile_murder_rate, select= c(UrbanPop, Murder))
```

