---
title: "Assigment_03"
author: "Hasan Palaz"
date: "October 21, 2017"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=150)
```




## ASSIGNMENT#3
In this homework, you will do some data analysis using R for the Forest Fire Data described
http://archive.ics.uci.edu/ml/machine-learning-databases/forest-fires/forestfires.csv
The dataset is used to find the relationship between the burned area of forest fires
and meteorological data.

######Q1
1. Import the data into R.
a. How many observations are there in the dataset?

b. How many observations are there with a fire (i.e., area>0)

c. How many observations are there with a rain (i.e., rain>0)

d. How many observations are there with both a fire and a rain?

######A1
1. Import the data into R.
```{r}
forestfires_df <- read.csv("C:\\Users\\Enespc\\Desktop\\HASAN\\Introduction to Data Analysis\\forestfires.csv", header = TRUE)
head(forestfires_df)
#View(forestfires_df)
```
a. How many observations are there in the dataset?
```{r}
nrow(forestfires_df)
```
b. How many observations are there with a fire (i.e., area>0)
```{r}
area_fire <- subset(forestfires_df, area > 0)
nrow(area_fire)
```
c. How many observations are there with a rain (i.e., rain>0)
```{r}
area_rain <- subset(forestfires_df, rain > 0)
nrow(area_rain)
```
d. How many observations are there with both a fire and a rain?
```{r}
area_fire_and_rain <- subset(forestfires_df, area >0 & rain > 0)
nrow(area_fire_and_rain)
```



######Q2
Show the columns month, day, area of the all the observations.

######A2
```{r}
new_format_all <- subset(forestfires_df, select = c("month", "day", "area"))
new_format_all
```

######Q3
Show the columns month, day, area of the observations with a fire 

######A3
```{r}
new_format_fire <- subset(forestfires_df, area > 0, select = c("month", "day", "area"))
nrow(new_format_fire)
new_format_fire
```

######Q4
How large are the five largest fires (i.e., having largest area) 
a.What are the corresponding month, temp, RH, wind, rain, area?
b.Reorder factor levels of month to be from Jan to Dec.
Add one column to the data indicating whether a fire occurred for each observation
(True for area >0 and False for area ==0)

######A4

```{r}
largest_fires <- forestfires_df[order(-forestfires_df$area),]
largest_fires[1:5,]
```

a.What are the corresponding month, temp, RH, wind, rain, area?

```{r}
largest_fires <- subset(largest_fires, select = c("month", "temp", "RH", "wind", "rain", "area"))
largest_fires[1:5,]
```

b.Reorder factor levels of month to be from Jan to Dec.

Add one column to the data indicating whether a fire occurred for each observation
(True for area >0 and False for area ==0)

```{r}
forestfires_df$fire_occurred <- forestfires_df$area > 0
#head(forestfires_df)
month <- tolower(month.abb)
forestfires_df$month <- factor(forestfires_df$month, month, ordered=TRUE)
newww <- forestfires_df[order(forestfires_df$month), ]
newww
```

